Fusion Moves for Correlation Clustering

### 2014-11-12 paper/egpaper_for_review.tex
work on exp sec

For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contracted graph directly
depends on the number of connected components
of the intersection of the proposal solution and the current best solution.
%
In the extreme case, where each 
node is in a separate connnected component,
the fusion move is equivalent to solving the original problem.
In the other extreme, where the proposal has 
a single connected component, the current best solution will not change.
% 
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:fusion_move_a} can
be done fast enough, but  on the other side 
large enough to define a large internal polytope and therefore a powerful move.
%
To this end we suggest two proposal generators.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to get
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
We draw $n/2$ negative edges, and assign different seed to the endpoints
of each edge.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, \sigma_{ws})$
is added to the edge weighs.

@@

For \textbf{CGC}~\cite{beier_2014_cvpr} we used the public available implementation\footnote{\label{opengm-cvpr2014}{\url{github.com/opengm/opengm/tree/cgc-cvpr2014}}}
of Beier~\etal and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we use an modified version of
 the public available implementation in OpenGM\footref{opengm-master} of Kappes~\etal, as described in Sec.~\ref{sec:tunedMC}.
%
%We realize that their implementation has some limitations on large problems and make the following modifications.
%Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
%Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
%during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
%This procedure is one magnitude faster, but might cause a few extra outer iterations.
%In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
%For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}.
We use the public available code of Micha Elsner\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
While this implementation uses full adjacency matrixes it does not scale as well as the algorithm can and therefor can not be applied on all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.

As reported in~\cite{beier_2014_cvpr}, CGC is faster and gives better energies then PlanarCC~\cite{yarkony_2012_eccv} and Expand \& Explore~\cite{bagon_2011_arxiv}.
Therefore we exclude those in our experiments.

We compare this methods to the following methods suggested in the present paper.
%
Energy Based Based Hierarchical Clustering (\textbf{EHC})
as described in sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC denoted by \textbf{EHC-CGC}.
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers denoted by 
\textbf{CCFusion-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}, respectively.
We set the size of the proposals fix to $10\%$ of the original problem size
and the standard derivation of the random edge noise to $1.5$.

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
 \begin{figure}[t]
    \centering
    \begin{tikzpicture}
    \begin{semilogxaxis}[  mark size=1pt,
     height= 4cm, 
    restrict y to domain=-80000:1000,
   %   restrict x to domain=10:4000,
    xlabel = {runtime},
    xmin = 10,
    xmax = 4100,
    ymax = 0,
    ymin = -80000,
    legend to name = ledgendPositionMC,
    legend columns=4,
    width = 0.9\columnwidth,
    scaled ticks = false
    ] 
    \addplot[color=red,mark=x] table[x=time, y=MC-CCFDB-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-old}
    \addplot[color=orange,mark=o] table[x=time, y=MC-CCFDB-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-new}
    \addplot[color=blue,mark=x] table[x=time, y=MC-CCIFD-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-old}
    \addplot[color=cyan,mark=o] table[x=time, y=MC-CCIFD-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-new}

    \end{semilogxaxis}
    \end{tikzpicture}
  \begin{center}
  \hypersetup{linkcolor = black}
  \ref{ledgendPositionMC}
  \hypersetup{linkcolor = red}
  \end{center}
    \caption{Comparison of the Multicut implementation of OpenGM and 
    our modified implementation, which can deal better with large scale problems. 
    However, for large scale problems it still does not scale.}
    \label{fig:at:knott-450b}
  \end{figure}
When using the  public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we realize that their implementation has some limitations on large problems.
This results in a very slow separation and make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.

\subsection{Parameter Choice for CCFusion}
Beside the choice of the proposal generator and fusion method CCFusion has some more parameters, which need to be set.
The most crucial one is the number of segments in the proposal. For HC we also have to set the noise which is use to generate diversity.
Fig.~\ref{fig:parameterchoice} shows a evaluation of the impact of this parameters for a single instance of knott-3d-450 averaged over several random seeds.
The runtime many depend on the number of clusters in the proposal (Fig.~\ref{fig:parameterchoice} left), which controls the size of the auxiliary move problems.
The level of noise has no major impact on the runtime.
The energy of the final solution improves with finer proposals since this increase the search-space of the moves.
Although the level of noise has to be large enough to generate diverse proposals but also not to large to generate proposals with low energy.
As shown in Fig.~\ref{fig:parameterchoice} right the useful parameter set is quite large. This allows us to use the same parameters for \emph{all}
experiments. However we would like to note, that in practice we can improve the performance by adjust these parameters for the specific problem setting.
%
%
\begin{figure}[th]
\centering
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-time2.pdf}}
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-value2.pdf}}
\caption{Empirical evaluation of the impact of noise used for proposal generation and proposal size.
  Proposals with many segments causes longer runtime. Noise seemed not to be a critical parameter but should be selected large enough.
}
\label{fig:parameterchoice}
\end{figure}

\subsection{Social Networks and Network Clustering}
A application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $1$. 
%
The first network is called \emph{Epinions}. 
This is who-trust-whom online social network of a a general consumer review site Epinions.com. 
Each directed edge $a\to b$ indicated that user $a$ trusts  or does not trust user $b$ is the edge-weight is positive or negative, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The second network is called \emph{Slashdot}. 
Slashdot is a technology-related news website know for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friends or foes. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges from which $76.73\%$ are positively weighted.

@@

As another example for network clustering we use the modularity-clustering models from~\cite{kappes_2014_benchmark_arxiv} which are small but fully connected.

>>>>

For correlation clustering fusion we add a third
property: \emph{size}.
The size of the contracted graph directly
depends on the number of connected components
of the intersection of the proposal solution and the current best solution.
%
In the extreme case, where each 
node is in a separate connnected component,
the fusion move is equivalent to solving the original problem.
In the other extreme, where the proposal has 
a single connected component, the current best solution will not change.
% 
Therefore the size of the proposals
should be small enough, such that solving
eq.~\ref{eq:fusion_move_a} can
be done fast enough, but  on the other side 
large enough to define a large internal polytope and so a powerful move.
%
To this end we suggest two proposal generators.

@@

\noindent \textbf{Randomized Watersheds (RWS):}
%\subsubsection{Randomized Watersheds (RWS)}
Watersheds have become quite popular for
graph segmentation and have a strong connection
to energy minimization~\cite{couprie_2011}.
%
The edge weighted watershed algorithm~\cite{meyer_2013}
with random seeds can be used to get
cheap proposals.
To improve \emph{quality} we do not use $n$ seeds distributed uniformly
over all nodes but use the following.
%
We draw $n/2$ negative edges, and assign to the endpoints a different seed.
Doing so, a random subset of negative edges is forced
to be cut within each proposal.
For additional \emph{diversity}, noise $\mathcal{N} (0, \sigma_{ws})$
can be added to the edge weighs.

@@

For \textbf{CGC}~\cite{beier_2014_cvpr} we used the public available implementation\footnote{\label{opengm-cvpr2014}{\url{github.com/opengm/opengm/tree/cgc-cvpr2014}}}
of the Beier~\etal and for \textbf{KL}~\cite{Kernighan-1970} the implementation in OpenGM\footlabel{opengm-master}{\url{github.com/opengm/opengm}}.
For integer multicuts (\textbf{MC-I}) and relaxed multicuts (\textbf{MC-R})~\cite{kappes_2013_arxiv} we use an modified version of
 the public available implementation in OpenGM\footref{opengm-master} of Kappes~\etal, as described in Sec.~\ref{sec:tunedMC}.
%
%We realize that their implementation has some limitations on large problems and make the following modifications.
%Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
%Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
%during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
%This procedure is one magnitude faster, but might cause a few extra outer iterations.
%In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
%For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
From the field of data-mining we compare to the PIVOT-algorithm~\cite{Ailon-2008} followed by a round of BOEM~\cite{Gionis-2007} denoted by \textbf{PIVOT-BOEM}.
We use the public available code of Micha Elsner\footnote{\url{http://www.ling.ohio-state.edu/~melsner/resources/correlation-readme.html}}.
While this implementation uses full adjacency matrixes it does not scale as well and can not be applied on all datasets.
We also run classical fusion moves~\cite{Lempitsky-2010} (\textbf{Fusion}) and select distinct labels for the two candidate segmentations.
%
As reported in~\cite{beier_2014_cvpr}, CGC is faster and gives better energies then PlanarCC~\cite{yarkony_2012_eccv} and Expand \& Explore~\cite{bagon_2011_arxiv}.
Therefore we exclude those in our experiments.

We compare this methods to the following methods suggested in the present paper.
%
Energy Based Based Hierarchical Clustering (\textbf{EHC}),
as described in Sec.~\ref{sec:ehc}.
%
CGC warm started with the solution from EHC denoted by \textbf{EHC-CGC}.
%
The proposed correlation cluster fusion algorithm with EHC-based and watershed-based proposals
 and MC-I and CGC as subproblem solvers denoted by 
\textbf{CCFusion-HC-MC},
\textbf{-HC-CGC},
\textbf{-WS-MC}, and 
\textbf{-WS-CGC}, respectively.
We set the size of the proposals fix to $10\%$ of the original problem size
and the standard derivation of the random edge noise is set to $1.5$.

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
When using the  public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we realize that their implementation has some limitations on large problems.
This results in a very slow separation and make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.

 \begin{figure}[t]
    \centering
    \begin{tikzpicture}
    \begin{semilogxaxis}[  mark size=1pt,
     height= 4cm, 
    restrict y to domain=-80000:1000,
   %   restrict x to domain=10:4000,
    xlabel = {runtime},
    xmin = 10,
    xmax = 4100,
    ymax = 0,
    ymin = -80000,
    legend to name = ledgendPositionMC,
    legend columns=4,
    width = 0.9\columnwidth,
    scaled ticks = false
    ] 
    \addplot[color=red,mark=x] table[x=time, y=MC-CCFDB-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-old}
    \addplot[color=orange,mark=o] table[x=time, y=MC-CCFDB-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCR-new}
    \addplot[color=blue,mark=x] table[x=time, y=MC-CCIFD-o*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-old}
    \addplot[color=cyan,mark=o] table[x=time, y=MC-CCIFD-n*]{data3/knott-3d-450b.data};          \addlegendentry{MCI-new}

    \end{semilogxaxis}
    \end{tikzpicture}
  \begin{center}
  \hypersetup{linkcolor = black}
  \ref{ledgendPositionMC}
  \hypersetup{linkcolor = red}
  \end{center}
    \caption{Comparison of the Multicut implementation of OpenGM and 
    our modified implementation, which can deal better with large scale problems. 
    However, for large scale problems it still does not scale.}
    \label{fig:at:knott-450b}
  \end{figure}

\subsection{Parameter Choice for CCFusion}
Beside the choice of the proposal generator and fusion method CCFusion has some more parameters, which need to be set.
The most crucial one is the number of segments in the proposal. For HC we also have to set the noise which is use to generate diversity.
Fig.~\ref{fig:parameterchoice} shows a evaluation of the impact of this parameters for a single instance of knott-3d-450 averaged over several random seeds.
The runtime many depend on the number of clusters in the proposal (Fig.~\ref{fig:parameterchoice} left), which controls the size of the auxiliary move problems.
The level of noise has no major impact on the runtime.
The energy of the final solution improves with finer proposals since this increase the search-space of the moves.
Although the level of noise has to be large enough to generate diverse proposals but also not to large to generate proposals with low quality.
As shown in Fig.~\ref{fig:parameterchoice} right the useful parameter set is quite large. This allows us to use the same parameters for \emph{all}
experiments. However we would like to note, that in practice we can improve the performance by adjust these parameters for the specific problem setting.
%
%
\begin{figure}[th]
\centering
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-time2.pdf}}
\fbox{\includegraphics[width=0.45\columnwidth,trim=3.5cm 8cm 4.5cm 8cm,clip]{images/grid-value2.pdf}}
\caption{Empirical evaluation of the impact of noise used for proposal generation and proposal size.
  Proposals with many segments causes longer runtime. Noise seemed not to be a critical parameter but should be selected large enough.
}
\label{fig:parameterchoice}
\end{figure}

\subsection{Social Networks and Network Clustering}
An application for large scale correlation clustering are social networks.
We consider two of those networks from the Stanford Large Network Dataset Collection\footnote{\url{http://snap.stanford.edu/data/index.html}}.
Both networks are given by weighted directed graphs with edge weights $-1$ and $1$. 
%
The first network is called \emph{Epinions}. 
This is who-trust-whom online social network of a a general consumer review site Epinions.com. 
Each directed edge $a\to b$ indicated that user $a$ trusts  or does not trust user $b$ by a  positive or negative edge-weight, respectively.
The network contains $131828$ nodes and $841372$ edges from which $85.3\%$ are positively weighted.
%
The second network is called \emph{Slashdot}. 
Slashdot is a technology-related news website know for its specific user community. 
In 2002 Slashdot introduced the Slashdot Zoo feature which allows users to tag each other as friends or foes. 
The network was obtained in November 2008 and contains $77350$ nodes and $516575$ edges from which $76.73\%$ are positively weighted.

@@

As shown in Fig.~\ref{fig:anytime} (g,h) CCFusion provides for the first minutes the best results.
Only CGC and HC-CGC are able to find better solutions after more then 1000 seconds.
MC-I and MC-R can even not be applied on such large problems.
We belief that with better proposal generators, which are more suited for such network problems,
we can improve CCFusion. One candidate for such a generated would be a scalable method for the 
PIVOT algorithm.

As another example for network clustering we use the modularity-clustering models from~\cite{kappes_2014_benchmark_arxiv} which are small but fully connected.
Although, CCFusion and the used parameters was not designed and chosen for this type of problem, it performs on par to competitive methods. 
Especially is does a better job than MC-I as shown in Fig.~\ref{fig:anytime} (i)
